---
output: github_document
editor_options:
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "www/README-",
  warning = FALSE,
  message = FALSE
)
```

[![Travis-CI Build Status](https://travis-ci.org/Weecology/NeonTreeEvaluation_package.svg?branch=master)](https://travis-ci.org/Weecology/NeonTreeEvaluation_package)

# A multi-sensor benchmark dataset for detecting individual trees in airborne RGB, Hyperspectral and LIDAR point clouds

Maintainer: Ben Weinstein - University of Florida.

This benchmark dataset is the first dataset to have consistent annotation approach across a variety of ecosystems. This repo is the R package for reproducible evaluation against the dataset. The benchmark dataset contains over 6,000 image-annotated crowns, 424 field-annotated crowns, and 3,777 overstory stem points from a wide range of forest types. 

# Installation

```{r,eval=F}
library(devtools)
install_github("Weecology/NeonTreeEvaluation_package")
```

# Download sensor data

To download evaluation data from the Zenodo archive (1GB), use the download() function to place the data in the correct package location. Download the much larger training data, set training=TRUE.

```{r, eval=FALSE}
library(NeonTreeEvaluation)
download()
```

# Getting Started

The package contains two vignettes. The 'Data' vignette describes each datatype and how to interact with it in R. The 'Evaluation' vignette shows how to submit predictions to the benchmark.

# Submission Format

## CSV bounding boxes

The format of the submission is as follows

* A csv file
* 5 columns: plot_name, xmin, ymin, xmax, ymax

Each row contains information for one predicted bounding box.

The plot_name should be named the same as the files in the dataset without extension (e.g. SJER_021 not SJER_021.tif) and not the full path to the file on disk. Not all evaluation data are available for all plots. Functions like evaluate_field_crowns and evaluate_image_crowns will look for matching plot name and ignore other plots.Depending on the speed of the algorithm, the simplest thing to do is predict all images in the RGB folder (see list_rgb()) and the package will handle matching images with the correct data to the correct evaluation procedure.

For a list of NEON site abbreviations: https://www.neonscience.org/field-sites/field-sites-map

### Example

The package contains a sample submission file. 

```{r}
library(raster)
library(dplyr)
library(NeonTreeEvaluation)
head(submission)
```

## Shp Polygons

Instead of bounding boxes, some methods may return polygons. To submit as polygons, create a single unprojected shapefile with polygons in image coordinates. Polygons must be complete with no holes. Here is an example of the above csv file in polygon format. Here the xmin, xmax, etc. columns are ignored since the information is stored in the geometry data.

```{r}
head(submission_polygons)
```

## Scores for an image-annotated crowns

| Author                | Precision | Recall | Cite/Code                              |
|-----------------------|-----------|--------|------------------------------------------|
| Weinstein et al. 2020 | 0.66 | 0.79 | https://deepforest.readthedocs.io/|
| Silva et al. 2016     | 0.34| 0.47 |[lidR package](https://www.rdocumentation.org/packages/lidR/versions/1.6.1/topics/lastrees_silva)|

The main data source are image-annotated crowns, in which a single observer annotated visible trees in 200 40m x 40m images from across the United States. This submission has bounding boxes in image coordinates. To get the benchmark score image-annotated ground truth data.

```{r}
#Get a three sample plots to run quickly, ignore to run the entire dataset
df<-submission %>% filter(plot_name %in% c("SJER_052"))

#Compute total recall and precision for the overlap data
results<-evaluate_image_crowns(predictions = df,project = T, show=F, summarize = T)
results[1:3]
```

For a list of NEON site abbreviations: https://www.neonscience.org/field-sites/field-sites-map

## Scores for an field-annotated crowns

| Author                | Recall | Cite/Code|
|-----------------------|--------|----------|
| Weinstein et al. 2020 | 0.61   |https://deepforest.readthedocs.io/ |

The second data source is a small number of field-annotated crowns from two geographic sites. These crowns were drawn on a tablet while physically standing in the field, thereby reducing the uncertainty in crown segmentation.

```{r}
df <- submission %>% filter(plot_name=="OSBS_95_competition")
results<-evaluate_field_crowns(predictions = df,project = T)
results
```


## Scores for an field-collected stems

| Author                |Recall | Cite/Code                              |
|-----------------------|--------|------------------------------------------|
| Weinstein et al. 2020 | 0.74 | https://deepforest.readthedocs.io/|

The third data source is the NEON Woody Vegetation Structure Dataset. Each tree stem is represented by a single point. This data has been filtered to represent overstory trees visible in the remote sensing imagery.

```{r}
df <- submission %>% filter(plot_name=="JERC_049")
results<-evaluate_field_stems(predictions = df,project = F, show=T, summarize = T)
results
```

If you would prefer not to clone this repo, a static version of the benchmark is here: https://zenodo.org/record/3723357#.XqT_HlNKjOQ

# Sensor Data

## RGB Camera

```{r}
library(raster)
library(NeonTreeEvaluation)

#Read RGB image as projected raster
rgb_path<-get_data(plot_name = "SJER_021",type="rgb")
rgb<-stack(rgb_path)

#Find path and parse
xmls<-get_data("SJER_021",type="annotations")
annotations<-xml_parse(xmls)
#View one plot's annotations as polygons, project into UTM
#copy project utm zone (epsg), xml has no native projection metadata
xml_polygons <- boxes_to_spatial_polygons(annotations,rgb)

plotRGB(rgb)
plot(xml_polygons,add=T)
```

## Lidar

To access the draped lidar hand annotations, use the "label" column. Each tree has a unique integer.

```{r}
library(lidR)
path<-get_data("TEAK_052",type="lidar")
r<-readLAS(path)
trees<-lasfilter(r,!label==0)
plot(trees,color="label")
```

We elected to keep all points, regardless of whether they correspond to tree annotation. Non-tree points have value 0. We highly recommend removing these points before predicting the point cloud. Since the annotations were made in the RGB and then draped on to the point cloud, there will naturally be some erroneous points at the borders of trees.

## Hyperspectral

Hyperspectral surface reflectance (NEON ID: DP1.30006.001) is a 426 band raster covering visible and near infrared spectrum.

```{r}
path<-get_data("MLBS_071",type="hyperspectral")
g<-stack(path)
nlayers(g)
#Grab a three band combination to view as false color
f<-g[[c(52,88,117)]]
plotRGB(f,stretch="lin")
```

# Submission Ranks

To add score to this benchmark, please submit a pull request to this README with the scores and the submission csv for confirmation.

## Citation
This benchmark is currently in review. Either cite this repo, or the original article using these data:
<sup>1</sup> Weinstein, Ben G., et al. "Individual tree-crown detection in RGB imagery using semi-supervised deep learning neural networks." Remote Sensing 11.11 (2019): 1309. https://www.mdpi.com/2072-4292/11/11/1309
