---
output:
  md_document:
    variant: markdown_github
editor_options: 
  chunk_output_type: console
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "www/README-",
  warning = FALSE,
  message = FALSE
)
```

[![Travis-CI Build Status](https://travis-ci.org/Weecology/NeonTreeEvaluation_package.svg?branch=master)](https://travis-ci.org/Weecology/NeonTreeEvaluation_package)

# A multi-sensor benchmark dataset for detecting individual trees in airborne RGB, Hyperspectral and LIDAR point clouds

Mantainer: Ben Weinstein - University of Florida.

Individual tree detection is a central task in forestry and ecology. Few papers analyze proposed methods across a wide geographic area. This limits the utility of tools and inhibits comparisons across methods. This benchmark dataset is the first dataset to have consistant annotation approach across a variety of ecosystems. 

This repo is the accomponaying R package for reproducible evaluation against the dataset. The package contains a static version of the annotations used for evaluation. 

# Installation

```{r,eval=F}
library(devtools)
install_github("https://github.com/weecology/NeonTreeEvaluation_package.git")
```

# Getting Started

The package contains two vignettes. The 'Data' vignette describes each datatype and how to interact with it in R. The 'Evaluation' vignette shows how to submit predictions to the benchmark.  

# Submission Format

The format of the submission is as follows

* A csv file
* 5 columns: Plot Name, xmin, ymin, xmax, ymax

Each row contains information for one predicted bounding box.

The plot column should be named the same as the files in the dataset (e.g. SJER_021), not the path to the file.

# Example

```{r}
library(raster)
library(dplyr)
library(NeonTreeEvaluation)
submission<-read.csv(system.file("extdata", "Weinstein2019.csv",package = "NeonTreeEvaluation"))
head(submission)
```

## Precision and recall scores for a single hand-annotated plot

This submission has bounding boxes in image coordinates. 
```{r}
df<-submission %>% filter(plot_name=="SJER_052")
evaluate_plot(df,show = T,project_boxes = T)
```

If you would prefer not to clone this repo, a static version of the benchmark is here: https://zenodo.org/record/3723357#.XqT_HlNKjOQ

### For the entire benchmark

```{r,eval=F}
evaluate_benchmark(submission,project_boxes = T)
```

# Sensor Data

The NeonTreeEvaluation dataset is a set of bounding boxes drawn on RGB imagery from the National Ecological Observation Network (NEON). NEON is a set of 45 sites (e.g. [TEAK](https://www.neonscience.org/field-sites/field-sites-map/TEAK)) that cover the dominant ecosystems in the US.

For more complete information see:

https://github.com/weecology/NeonTreeEvaluation

## RGB

```{r}
library(raster)
library(NeonTreeEvaluation)

#Read RGB image as projected raster
rgb_path<-get_data(plot_name = "SJER_021",type="rgb")
rgb<-stack(rgb_path)

#Path to dataset
xmls<-readTreeXML(siteID="SJER")

#View one plot's annotations as polygons, project into UTM
#copy project utm zone (epsg), xml has no native projection metadata
xml_polygons <- boxes_to_spatial_polygons(xmls[xmls$filename %in% "SJER_021.tif",],rgb)

plotRGB(rgb)
plot(xml_polygons,add=T)
```

## Lidar

To access the draped lidar hand annotations, use the "label" column. Each tree has a unique integer.

```{r}
library(lidR)
path<-get_data("TEAK_052",type="lidar")
r<-readLAS(path)
trees<-lasfilter(r,!label==0)
plot(trees,color="label")
```

We elected to keep all points, regardless of whether they correspond to tree annotation. Non-tree points have value 0. We highly recommend removing these points before predicting the point cloud. Since the annotations were made in the RGB and then draped on to the point cloud, there will naturally be some erroneous points at the borders of trees.

## Hyperspectral 

Hyperspectral surface reflectance (NEON ID: DP1.30006.001) is a 426 band raster covering visible and near infared spectrum.

```{r}
path<-get_data("MLBS_071",type="hyperspectral")
g<-stack(path)
nlayers(g)
#Grab a three band combination to view as false color
f<-g[[c(52,88,117)]]
plotRGB(f,stretch="lin")
```
 
# Submission

To submit to this benchmark, please see the evaluation vignette. Please submit a pull request, or contact the mantainer if you use these data in analysis and would like the results to be shown here.

## Citation
<sup>1</sup> Weinstein, Ben G., et al. "Individual tree-crown detection in RGB imagery using semi-supervised deep learning neural networks." Remote Sensing 11.11 (2019): 1309. https://www.mdpi.com/2072-4292/11/11/1309


