---
title: "Benchmark Evaluation"
author: "Ben Weinstein"
date: "`r Sys.Date()`"
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Submission Format

The format of the submission is as follows

* A csv file
* 5 columns: Plot Name, xmin, ymin, xmax, ymax

Each row contains information for one predicted bounding box.

The plot column should be named the same as the files in the dataset (e.g. SJER_021)

```{r}
library(dplyr)
library(NeonTreeEvaluation)
library(reshape2)
library(ggplot2)

Weinstein2019<-read.csv(system.file("extdata", "Weinstein2019.csv",package = "NeonTreeEvaluation"))

Silva2016<-read.csv(system.file("extdata", "Silva2016.csv",package = "NeonTreeEvaluation"))

Dalponte2016<-read.csv(system.file("extdata", "Dalponte2016.csv",package = "NeonTreeEvaluation"))

Li2012<-read.csv(system.file("extdata", "Li2012.csv",package = "NeonTreeEvaluation"))
```

# Evaluation

For each plot compute the precision and recall based on intersection-over-union of 0.5 between the ground truth bounding boxes and the predicted boxes. 

```{r}
#Run for each method.
Weinstein2019_results<-Weinstein2019 %>% group_by(plot_name)  %>% do(evaluate_plot(.,show=TRUE)) %>% mutate(Method="Weinstein2019")

Silva2016_results<-Silva2016 %>% group_by(plot_name)  %>%  do(evaluate_plot(.,project_boxes=FALSE,show=TRUE)) %>%  mutate(Method="Silva2016")

Dalponte2016_results<-Dalponte2016 %>% group_by(plot_name)  %>%  do(evaluate_plot(.,project_boxes=FALSE,show=TRUE)) %>%  mutate(Method="Dalponte2016")

Li2012_results<-Li2012 %>% group_by(plot_name)  %>%  do(evaluate_plot(.,project_boxes=FALSE,show=TRUE)) %>%  mutate(Method="Li2012")

results<-bind_rows(list(Weinstein2019_results,Silva2016_results,Dalponte2016_results,Li2012_results))
```

```{r}
df<-melt(results,id.vars=c("Method","plot_name"))
ggplot(df,aes(y=value,x=variable,fill=Method)) + geom_boxplot()
```


